---
title: "Comorbidities_Draft1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Reading the data
```{r }
rm(list = ls())

setwd("/rds/general/project/hda-22-23/live/TDS/Group3_Temporary/extraction_and_recoding/outputs")
ukb_extracted <- readRDS("/rds/general/project/hda-22-23/live/TDS/Group3_Temporary/extraction_and_recoding/outputs/ukb_extracted.rds")
output_final <- readRDS("/rds/general/project/hda-22-23/live/TDS/Group3_Temporary/outcome_definition/Outputs/output_final.rds")
```


# Combining the 2 dataframes

```{r }

participants = rownames(output_final)
output_final = cbind(participants, output_final)
participants2 = rownames(ukb_extracted)
ukb_extracted = cbind(participants2 , ukb_extracted)

joined_df <- output_final %>% 
  left_join(ukb_extracted, by = c("participants" = "participants2"))
dim (joined_df)
#View(head(joined_df))

joined_df = joined_df %>% select(-(participants))

#Removing prevalent cases 
prevalent_removed <- joined_df%>%
  filter(prevalent_case == 0)
nrow(prevalent_removed)


# the NAs were because of the prevalent cases
# Incidenct_NA_rmvd <- prevalent_removed%>%
#   filter(is.na(incident_case) != T)
# nrow(Incidenct_NA_rmvd)

df = prevalent_removed

```

# Loading the libraries
```{r }
library(table1)
library(dplyr)
library(hash)
library(knitr)
library(vtable)
library(tidyverse)
library(gtsummary)
library(ggplot2)
```

# Selecting the instances 0.0
```{r }
# variables = grep(pattern=".0.0",x=names(ukb_extracted),value=TRUE)
# selection = ukb_extracted[, variables]
# dim(selection)
# 
# # class(selection)
# View(head(selection))
# View (ukb_extracted)
# # View(selection)
```

# Checking the ICD10 codes

```{r }
# sum(selection$`Diagnoses_-_ICD10.0.0`=="")
# class(selection$`Diagnoses_-_ICD10.0.0`=="")
# head(selection$`Diagnoses_-_ICD10.0.0`=="")
# 
# subset2 = subset(selection, selection$`Diagnoses_-_ICD10.0.0`=="")
# 
# View(head(subset2))
# dim(subset2)
# 
# sum(selection$`Diagnoses_-_ICD10.0.242`=="")
# nrow(selection)
# 
# mycols <- subset2 %>% select(matches("ICD"))
# dim(mycols)
# View(head(mycols))
# 
# missing_vals <- mycols %>% summarise_all(funs(sum(. == "")))
# sum(missing_vals)
# View(missing_vals)
```


# Making a copy of the dataframe
```{r }
#df = selection[1:100000,]
# df = selection
# 
# dim(df)

```

# Creating a data frame for the ICD10, ICD9 and fields 20002, 20003 codes only

```{r }
# ICD10_DF
ICD10_cols =grep ("ICD10", names(df))
ICD10_data = df [, ICD10_cols]

#View(head(SR_data))

#ICD9_DF
ICD9_cols =grep ("ICD9", names(df))
ICD9_data = df [, ICD9_cols]

#SR DF
SR_cols =grep ("Non-cancer_illness_code,_self-reported", names(df))
SR_data = df[, SR_cols]

#Treatment and Medication DF
TM_cols =grep ("Treatment/medication_code_diag", names(df))
TM_data = df [, TM_cols]

#Illness of father DF
IF_cols =grep ("Illnesses_of_father", names(df))
IF_data = df [, IF_cols]

#Illness of mother DF
IM_cols =grep ("Illnesses_of_mother", names(df))
IM_data =df [, IM_cols]

#Illness of siblings DF
IS_cols =grep ("Illnesses_of_siblings", names(df))
IS_data = df [, IS_cols]

#Medication for cholesterol, blood pressure, or diabetes DF
MCBP_cols =grep ("Medication_for_cholesterol,_blood_pressure_or_diabetes", names(df))
MCBP_data = df [, MCBP_cols]

#Medication for cholesterol, blood pressure, or diabetes or take exogenous hormones DF
MCBPEH_cols =grep ("Medication_for_cholesterol,_blood_pressure_or_diabetes", names(df))
MCBPEH_data = df [, MCBPEH_cols]

# Operation codes

OPCS_cols =grep ("Operative_procedures", names(df))
OPCS_data = df [, OPCS_cols]


#Creating a dataframe that will containe the variables of comorbidities
df2 = subset(df, select= c(eid, incident_case))



```


# Atrial Fibrillation

```{r }
# Create a list of ICD 10s to search for
AF_ICDs <- c("I48")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_AF_icd10 <- apply(ICD10_data, 1, function(x) {
  ifelse(any(grepl(paste(AF_ICDs, collapse = "|"), x)), 1, 0)
})
table(df$has_AF_icd10)


#Selecting from the ICD-9 column

# Atrial Fibrillation
# list of ICD-9 to match
ICD9_AF <- c("4273","4270")

# create has_AF column if they have the disease
df$has_AF_icd9 = apply(ICD9_data, 1, function(x) {
  ifelse(any(grepl(paste(ICD9_AF, collapse = "|"), x)), 1, 0)
})
table(df$has_AF_icd9)
# Selecting from The self-reported data-field from the data field 20002
SR_AF = c("1471","1483")
df$has_AF_SR = apply(SR_data, 1, function(x) {
  ifelse(any(grepl(paste(SR_AF, collapse = "|"), x)), 1, 0)
})


# Selecting from the OPCS-4
# Create a list of ICD 10s to search for
AF_OPCS <- c("K622", "K623")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_AF_opcs <- apply(OPCS_data, 1, function(x) {
  ifelse(any(grepl(paste(AF_OPCS, collapse = "|"), x)), 1, 0)
})





df2$has_AF <- rowSums(df[, c("has_AF_icd10", "has_AF_icd9", "has_AF_SR", "has_AF_opcs")])
df2$com_has_AF <- ifelse(df2$has_AF > 0, 1, df2$has_AF)
#View(df2)
#table(df2$incident_case, df2$has_AF)






```
# cpode for table
```{r }


table_AF <- vtable(df2,stats = c("Count", "Percent"),vars = c("incident_case", "has_AF"))
# Print the resulting table

print(table_AF)
contingency_table <- table(df2$incident_case, df2$has_AF)
table_AF <- as.table(contingency_table)
dimnames(table_AF) <- list(incident_case = c("0", "1"),
 has_AF = c("0", "1", "2", "3"))
table_AF
class(table_AF)
table_df <- as.data.frame(table_AF)

# Create ggplot table
table_plot <- ggplot(table_df, aes(x = incident_case, y = has_AF, fill = Count)) + 
  geom_tile() +  
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# Save table as PNG image
ggsave("table.png", table_plot, width = 8, height = 6, dpi = 300)
getwd()

# Create table with sum and percentage using tbl_summary for certain variables

table1 <- tbl_summary(df2[, c("incident_case", "has_AF")]) %>%


  add_n() %>%



  modify_header(n = "**Count**") 

 

# Print the resulting table

table1

```

# Migraine

```{r }
# Create a list of ICDs to search for
Migraine_ICDs <- c("G43", "G440", "N943")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_migraine_icd10 <- apply(ICD10_data, 1, function(x) {
  ifelse(any(grepl(paste(Migraine_ICDs, collapse = "|"), x)), 1, 0)
})


#Selecting from the ICD-9 column

# Migraine
# list of ICD-9 to match
ICD9_migraine <- c("346")

# create has_AF column if they have the disease
df$has_migraine_icd9 = apply(ICD9_data, 1, function(x) {
  ifelse(any(grepl(paste(ICD9_migraine, collapse = "|"), x)), 1, 0)
})

# Selecting from The self-reported data-field from the data field 20002
SR_migraine = c("1265")
df$has_migraine_SR = apply(SR_data, 1, function(x) {
  ifelse(any(grepl(paste(SR_migraine, collapse = "|"), x)), 1, 0)
})

table(df$has_migraine_icd10)
table(df$has_migraine_icd9)
table(df$has_migraine_SR)

df2$has_migraine <- rowSums(df[, c("has_migraine_icd10", "has_migraine_icd9", "has_migraine_SR")])
df2$com_has_migraine <- ifelse(df2$has_migraine > 0, 1, df2$has_migraine)
#View(df2)
#table(df2$incident_case, df2$has_migraine)


# table1 <- tbl_summary(df2[, c("incident_case", "has_migraine")]) %>%
#   add_n() %>%
#  modify_header(n = "**Count**") 
# 
# # Print the resulting table
# 
# table1
```
# Removing unneeded object
```{r }

rm(joined_df, AF_ICDs, AF_OPCS, ICD9_AF, ICD9_cols, ICD9_migraine, participants, participants2, SR_AF, SR_migraine)

```

# Type 1 diabetes

```{r }
# Create a list of ICDs to search for
t1d_ICDs <- c("E10", "O230")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_t1d_icd10 <- apply(ICD10_data, 1, function(x) {
  ifelse(any(grepl(paste(t1d_ICDs, collapse = "|"), x)), 1, 0)
})


#Selecting from the ICD-9 column

# Type 1 diabetes
# list of ICD-9 to match
ICD9_t1d <- c("25001", "25011", "25021", "25031", "25041", "25051",
              "25061", "25071", "25081", "25091", "25003", "25013", 
              "25023", "25033", "25043", "25053", "25063", "25073", 
              "25083", "25093")

# create has_t1d column if they have the disease
df$has_t1d_icd9 = apply(ICD9_data, 1, function(x) {
  ifelse(any(grepl(paste(ICD9_t1d, collapse = "|"), x)), 1, 0)
})

# Selecting from The self-reported data-field from the data field 20002
SR_t1d = c("1222")
df$has_t1d_SR = apply(SR_data, 1, function(x) {
  ifelse(any(grepl(paste(SR_t1d, collapse = "|"), x)), 1, 0)
})

table(df$has_t1d_icd10)
table(df$has_t1d_icd9)
table(df$has_t1d_SR)

df2$has_t1d <- rowSums(df[, c("has_t1d_icd10", "has_t1d_icd9", "has_t1d_SR")])
df2$com_has_t1d <- ifelse(df2$has_t1d > 0, 1, df2$has_t1d)
# View(df2)
# table(df2$incident_case, df2$has_t1d)
# table1 <- tbl_summary(df2[, c("incident_case", "has_t1d")]) %>%
#   add_n() %>%
#  modify_header(n = "**Count**") 
# 
# # Print the resulting table
# 
# table1


```



# Type 2 diabetes

```{r }
# list of ICD 10
t2d_codes <- c("E11", "O231")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_t2d_icd10 <- apply(ICD10_data, 1, function(x) {
  ifelse(any(grepl(paste(t2d_codes, collapse = "|"), x)), 1, 0)
})
# list of ICD-9 to match
ICD9_t2d <- c("25000", "25010", "25020", "25030", "25040", "25050",
              "25060", "25070", "25080", "25090", "25002", "25012", 
              "25022", "25032", "25042", "25052", "25062", "25072", 
              "25082", "25092")

# create has_t2d column if they have the disease
df$has_t2d_icd9 = apply(ICD9_data, 1, function(x) {
  ifelse(any(grepl(paste(ICD9_t2d, collapse = "|"), x)), 1, 0)
})

# Selecting from The self-reported data-field from the data field 20002
SR_t2d = c("1223","1220")
df$has_t2d_SR =  apply(SR_data, 1, function(x) {
  ifelse(any(grepl(paste(SR_t2d, collapse = "|"), x)), 1, 0)
})


#Selecting the treatment and medication code from the data field 20003
TM_t2d = c("1140868902", "1140874646", "1140874674", "1140874718", 
          "1140874744", "1140883066", "1140884600", "1141152590", 
          "1141157284", "1141168660", "1141171646", "1141173882", 
          "1141189090" )
df$has_t2d_TM = apply(TM_data, 1, function(x) {
  ifelse(any(grepl(paste(TM_t2d, collapse = "|"), x)), 1, 0)
})

# if in field 2443


df$has_t2d_2443 <- ifelse(df$Diabetes_diag_bio.0.0 ==1, 1,0)

#table(df$has_t2d_icd10)
table(df$has_t2d_icd9)
table(df$has_t2d_SR)
table(df$has_t2d_TM)
#table(df$has_t2d_2443)

df2$has_t2d <- rowSums(df[, c("has_t2d_icd10", "has_t2d_icd9", "has_t2d_SR", "has_t2d_TM","has_t2d_2443" )])
df2$com_has_t2d <- ifelse(df2$has_t2d > 0, 1, df2$has_t2d)

# View(df2)
# table(df2$incident_case, df2$has_t2d)
# 
# table1 <- tbl_summary(df2[, c("incident_case", "has_t2d")]) %>%
#   add_n() %>%
#  modify_header(n = "**Count**") 
# 
# # Print the resulting table
# 
# table1
saveRDS(df2, file = "comorbidities.rds")



```

```{r }
rm(ICD9_t1d, ICD9_t2d, ICD9_AF, ICD9_migraine, Migraine_ICDs, t1d_ICDs, t2d_codes)

```


# Chonic kidney disease

```{r }
#list of ICD 10
ckd_icd10_codes <- c("N183", "N184", "N185")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_ckd_icd10 <- apply(ICD10_data, 1, function(x) {
  ifelse(any(grepl(paste(ckd_icd10_codes, collapse = "|"), x)), 1, 0)
})

# list of ICD-9 to match
ICD9_ckd <- c("5853", "5954", "5855")

# create has_t2d column if they have the disease
df$has_ckd_icd9 = apply(ICD9_data, 1, function(x) {
  ifelse(any(grepl(paste(ICD9_ckd, collapse = "|"), x)), 1, 0)
})

# Selecting from The self-reported data-field from the data field 20002
SR_ckd = c("1192","1193")
df$has_ckd_SR =  apply(SR_data, 1, function(x) {
  ifelse(any(grepl(paste(SR_ckd, collapse = "|"), x)), 1, 0)
})


#table(df$has_ckd_icd10)
table(df$has_ckd_icd9)
table(df$has_ckd_SR)

df2$has_ckd <- rowSums(df[, c("has_ckd_icd10", "has_ckd_icd9", "has_ckd_SR" )])
df2$com_has_ckd <- ifelse(df2$has_ckd > 0, 1, df2$has_ckd)
# View(df2)
# table(df2$incident_case, df2$has_ckd)
# 
# table1 <- tbl_summary(df2[, c("incident_case", "has_ckd")]) %>%
#   add_n() %>%
#  modify_header(n = "**Count**") 
# 
# # Print the resulting table
# 
# table1

```



# Rheumathoid arthrisis

```{r }
# ICD10
ra_icd10_codes <- c("M05", "M06")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_ra_icd10 <- apply(ICD10_data, 1, function(x) {
  ifelse(any(grepl(paste(ra_icd10_codes, collapse = "|"), x)), 1, 0)
})

# list of ICD-9 to match
ICD9_RA <- c("714")

# create has_t2d column if they have the disease
df$has_RA_icd9 = apply(ICD9_data, 1, function(x) {
  ifelse(any(grepl(paste(ICD9_RA, collapse = "|"), x)), 1, 0)
})

# Selecting from The self-reported data-field from the data field 20002
SR_RA = c("1464")
df$has_RA_SR =  apply(SR_data, 1, function(x) {
  ifelse(any(grepl(paste(SR_RA, collapse = "|"), x)), 1, 0)
})



#table(df$has_RA_icd10)
table(df$has_RA_icd9)
table(df$has_RA_SR)



df2$has_RA<- rowSums(df[, c("has_ra_icd10", "has_RA_icd9", "has_RA_SR" )])
df2$com_has_RA <- ifelse(df2$has_RA > 0, 1, df2$has_RA)
#View(df2)
# table(df$incident_case, df2$has_RA)



# 
# table1 <- tbl_summary(df2[, c("incident_case", "has_RA")]) %>%
#   add_n() %>%
#  modify_header(n = "**Count**") 
# 
# # Print the resulting table
# 
# table1

```

# Systemic lupus erythematosus

```{r }
#ICD10
sle_icd10_codes <- c("M32")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_sle_icd10 <- apply(ICD10_data, 1, function(x) {
  ifelse(any(grepl(paste(sle_icd10_codes, collapse = "|"), x)), 1, 0)
})

# list of ICD-9 to match
ICD9_sle <- c("7100")

# create has_t2d column if they have the disease
df$has_sle_icd9 = apply(ICD9_data, 1, function(x) {
  ifelse(any(grepl(paste(ICD9_sle, collapse = "|"), x)), 1, 0)
})

# Selecting from The self-reported data-field from the data field 20002
SR_sle = c("1381")
df$has_sle_SR =  apply(SR_data, 1, function(x) {
  ifelse(any(grepl(paste(SR_sle, collapse = "|"), x)), 1, 0)
})



#table(df$has_sle_icd10)
table(df$has_sle_icd9)
table(df$has_sle_SR)
#

df2$has_sle <- rowSums(df[, c("has_sle_icd10", "has_sle_icd9", "has_sle_SR" )])
df2$com_has_sle <- ifelse(df2$has_sle > 0, 1, df2$has_sle)
# View(df2)
# table(df2$incident_case, df2$has_sle)
saveRDS(df2, file = "comorbidities.rds")
```
```{r }
rm(prevalent_removed, ckd_icd10_codes, ICD9_ckd,ICD9_RA, ICD9_sle, ra_icd10_codes, sle_icd10_codes, SR_ckd, SR_AF)

```

# Severe Mental Illness

```{r }
#ICD10
psychiatric_icd10_codes <- c("F03", "F068", "F09", "F20", 
                             "F22", "F23", "F259", "F28",
                             "F29", "F31", "F39", "F53", "F333")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_psychiatric_icd10 <- apply(ICD10_data, 1, function(x) {
  ifelse(any(grepl(paste(psychiatric_icd10_codes, collapse = "|"), x)), 1, 0)
})

# list of ICD-9 to match
ICD9_psychiatric <- c("295", "298", "296")

# create has_t2d column if they have the disease
df$has_psychiatric_icd9 = apply(ICD9_data, 1, function(x) {
  ifelse(any(grepl(paste(ICD9_psychiatric, collapse = "|"), x)), 1, 0)
})

# Selecting from The self-reported data-field from the data field 20002
SR_psychiatric = c("1289","1291")
df$has_psychiatric_SR =  apply(SR_data, 1, function(x) {
  ifelse(any(grepl(paste(SR_psychiatric, collapse = "|"), x)), 1, 0)
})



#table(df$has_psychiatric_icd10)
table(df$has_psychiatric_icd9)
table(df$has_psychiatric_SR)
#
df2$has_psychiatric <- rowSums(df[, c("has_psychiatric_icd10", "has_psychiatric_icd9", "has_psychiatric_SR" )])
df2$com_has_psychiatric <- ifelse(df2$has_psychiatric > 0, 1, df2$has_psychiatric)
#View(df2)
#table(df2$incident_case, df2$has_psychiatric)

```



# Erectile Dysfunction

```{r }
#ICD10

erectile_dysfunction_icd10_codes <- c("N484", "N52")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_ED_icd10 <- apply(ICD10_data, 1, function(x) {
  ifelse(any(grepl(paste(erectile_dysfunction_icd10_codes, collapse = "|"), x)), 1, 0)
})




# Erectile dysfunction ICD9
ICD9_ED <- c("60784")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_ED_icd9<- apply(ICD9_data, 1, function(x) {
  ifelse(any(grepl(paste(ICD9_ED, collapse = "|"), x)), 1, 0)
})


# Erectile dysfunction self reported
SR_ED<- c("1518")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_ED_SR <- apply(SR_data, 1, function(x) {
  ifelse(any(grepl(paste(SR_ED, collapse = "|"), x)), 1, 0)
})


#### 20003 medications`
# Erectile dysfunction

# Erectile dysfunction medications
erectile_dysfunction_meds_20003 <- c("1141168936",
"1141168948", "1141168944", "1141168946", "1140869100", "1140883010")

# Add a new column that represents whether a row contains any string from the list of fruits
df$has_ED_meds20003 <- apply(TM_data, 1, function(x) {
  ifelse(any(grepl(paste(erectile_dysfunction_meds_20003, collapse = "|"), x)), 1, 0)
})



df2$has_ED <- rowSums(df[, c("has_ED_icd10", "has_ED_icd9", "has_ED_SR", "has_ED_meds20003" )])
df2$com_has_ED <- ifelse(df2$has_ED > 0, 1, df2$has_ED)
# View(df2)
# table(df2$incident_case, df2$has_ED)

```

#CAD in a 1st degree relative

```{r }
#ILness in father
IF_CAD<- c("1")

# Add a new column that represents whether a row contains any string from the list of fruits
df$IF_CAD <- apply(IF_data, 1, function(x) {
  ifelse(any(grepl(paste(IF_CAD, collapse = "|"), x)), 1, 0)
})

#ILness in mother
IM_CAD<- c("1")

# Add a new column that represents whether a row contains any string from the list of fruits
df$IM_CAD <- apply(IM_data, 1, function(x) {
  ifelse(any(grepl(paste(IM_CAD, collapse = "|"), x)), 1, 0)
})

#ILness in SIBLINGS
IS_CAD<- c("1")

# Add a new column that represents whether a row contains any string from the list of fruits
df$IS_CAD <- apply(IS_data, 1, function(x) {
  ifelse(any(grepl(paste(IS_CAD, collapse = "|"), x)), 1, 0)
})

df2$herit_CAD <- rowSums(df[, c("IF_CAD", "IM_CAD", "IS_CAD")])
df2$com_herit_CAD <- ifelse(df2$herit_CAD > 0, 1, df2$herit_CAD)
# View(df2)
# table(df2$incident_case, df2$herit_CAD)
saveRDS(df2, file = "comorbidities.rds")

```
# Atypical antipsychotic

```{r }
###### Atypical antipsychotic medication
antipsychotic_meds_20003 <- c("1140867420", "1140867444", "1140927956",
                              "1140928916", "1141152848", "1141153490",
                              "1141169714", "1141195974")

# Add a new column that represents whether a row contains any string from the list of fruits
df$antipsychotic_meds_20003 <- apply(TM_data, 1, function(x) {
  ifelse(any(grepl(paste(antipsychotic_meds_20003, collapse = "|"), x)), 1, 0)
})

table(df$antipsychotic_meds_20003)

df2$antipsychotic <- df$antipsychotic_meds_20003
df2$com_antipsychotic <- ifelse(df2$antipsychotic > 0, 1, df2$antipsychotic)
# View(df2)
# table(df2$incident_case, df2$antipsychotic)

```

# steroids meds
```{r }
steroid_meds_20003 <- c("1140874790", "1140874816", "1140874896.00",
                        "1140874930", "1140874976", "1141145782", 
                        "1141173346")

# Add a new column that represents whether a row contains any string from the list of fruits
df$steroid_meds_20003 <- apply(TM_data, 1, function(x) {
  ifelse(any(grepl(paste(steroid_meds_20003, collapse = "|"), x)), 1, 0)
})

table(df$steroid_meds_20003)

df2$com_steroid <- df$steroid_meds_20003

# View(df2)
# table(df2$incident_case, df2$steroid)
saveRDS(df2, file = "comorbidities.rds")

```


# Lipid lowering meds

```{r }

#  Lipid Lowering meds
lipid_lowering_meds_20003 <- c("1140861954",
                              "1140861958", "1140888594",
                              "1140888648", "1141146234", 
                              "1141192410", "1141192736")

# Add a new column that represents whether a row contains any string from the list of fruits
df$lipid_lowering_meds_20003 <- apply(TM_data, 1, function(x) {
  ifelse(any(grepl(paste(lipid_lowering_meds_20003, collapse = "|"), x)), 1, 0)
})

table(df$lipid_lowering_meds_20003)

df2$com_lipid_lowering <- df$lipid_lowering_meds_20003
# View(df2)
# table(df2$incident_case, df2$lipid_lowering)

```


# Blood pressure lowering medication
```{r }

bp_meds_20003 <- c("1140860192", "1140860292", "1140860696", "1140860728",
                   "1140860750", "1140860806", "1140860882", "1140860904", 
                   "1140861088", "1140861190", "1140861276", "1140866072", 
                   "1140866078", "1140866090", "1140866102", "1140866108",
                   "1140866122", "1140866138", "1140866156", "1140866162",
                   "1140866724", "1140866738", "1140868618", "1140872568",
                   "1140874706", "1140874744", "1140875808", "1140879758", 
                   "1140879760", "1140879762", "1140879802", "1140879806",
                   "1140879810", "1140879818", "1140879822", "1140879826",
                   "1140879830", "1140879834", "1140879842", "1140879866", 
                   "1140884298", "1140888552", "1140888556", "1140888560", 
                   "1140888646", "1140909706", "1140910442", "1140910614", 
                   "1140916356", "1140923272", "1140923336", "1140923404", 
                   "1140923712", "1140926778", "1140928226", "1141145660", 
                   "1141146126", "1141152998", "1141153026", "1141164276", 
                   "1141165470", "1141166006", "1141169516", "1141171336", 
                   "1141180592", "1141180772", "1141180778", "1141184722", 
                   "1141193282", "1141194794", "1141194810")

# Add a new column that represents whether a row contains any string from the list of fruits
df$bp_meds_20003 <- apply(TM_data, 1, function(x) {
  ifelse(any(grepl(paste(bp_meds_20003, collapse = "|"), x)), 1, 0)
})

table(df$bp_meds_20003)

# if in field 6177

MCBP_bp_meds<- c("2")

# Add a new column that represents whether a row contains any string from the list of fruits
df$MCBP_bp_meds_6177 <- apply(MCBP_data, 1, function(x) {
  ifelse(any(grepl(paste(MCBP_bp_meds, collapse = "|"), x)), 1, 0)
})
# if in field 6153

BPMC_bp_meds<- c("2")

# Add a new column that represents whether a row contains any string from the list of fruits
df$MCBPEH_bp_meds_6153 <- apply(MCBPEH_data, 1, function(x) {
  ifelse(any(grepl(paste(BPMC_bp_meds, collapse = "|"), x)), 1, 0)
})

df2$bp = rowSums(df[, c("bp_meds_20003", "MCBP_bp_meds_6177", "MCBPEH_bp_meds_6153" )])
df2$com_bp <- ifelse(df2$bp > 0, 1, df2$bp)

# View(df2)
# table(df2$incident_case, df2$bp)

saveRDS(df2, file = "comorbidities.rds")
getwd()

```


# WE ADDED 1 IN BP because 1 refers to lipid lowering medications!!!!
Elliot J, Bodinier B, Bond TA, et al. Predictive accuracy of a polygenic risk score–
enhanced prediction model vs a clinical risk score for coronary artery disease. 
JAMA. doi:10.1001/jama.2019.22241 

