---
title: "Data For Imputation"

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Loading the libraries

```{r }
rm(list = ls())
library(tableone)
library(table1)
library(dplyr)
library(tidyr)
library(pheatmap)
library(RColorBrewer)
library(pheatmap)


```

# Reading the data

```{r }

setwd("/rds/general/project/hda-22-23/live/TDS/Group3_Temporary/extraction_and_recoding/outputs")
ukb_extracted <- readRDS("/rds/general/project/hda-22-23/live/TDS/Group3_Temporary/extraction_and_recoding/outputs/ukb_extracted.rds")
output_final <- readRDS("/rds/general/project/hda-22-23/live/TDS/Group3_Temporary/outcome_definition/Outputs/output_final.rds")
comorbidities <- readRDS("/rds/general/project/hda-22-23/live/TDS/Group3_Temporary/FINAL_SCRIPTS/comorbidities.rds")

prs_69328 <- readRDS("/rds/general/user/rw1317/projects/hda-22-23/live/TDS/General/Data/prs_69328.rds")


```


# Combining the 2 dataframes and adding the PRS column

```{r }

participants = rownames(output_final)
output_final = cbind(participants, output_final)
participants2 = rownames(ukb_extracted)
ukb_extracted2 = cbind(participants2 , ukb_extracted)

joined_df <- output_final %>% 
  left_join(ukb_extracted2, by = c("participants" = "participants2"))
dim (joined_df)
#View(head(joined_df))

joined_df = joined_df %>% select(-(participants))

#Removing prevalent cases 
prevalent_removed <- joined_df%>%
  filter(prevalent_case == 0)
nrow(prevalent_removed)


# the NAs were because of the prevalent cases
Incidenct_NA_rmvd <- prevalent_removed%>%
  filter(is.na(incident_case) != T)
nrow(Incidenct_NA_rmvd)

df = prevalent_removed
rm(Incidenct_NA_rmvd, prevalent_removed, ukb_extracted, ukb_extracted2, participants, participants2, output_final, joined_df)

polygenic_score = data.frame(eid = names(prs_69328), PRS = prs_69328, row.names = NULL)
df <- merge(df , polygenic_score, by = "eid")


```

# Removing the variables associated with the comorbidities 

```{r }

df <- df %>%
  select(-which(sapply(names(df), grepl, pattern = "Operative_procedures")))


df <- df %>%
  select(-which(sapply(names(df), grepl, pattern = "ICD10")))


df <- df %>%
  select(-which(sapply(names(df), grepl, pattern = "ICD9")))

df <- df %>%
  select(-which(sapply(names(df), grepl, pattern = "Cancer_code,_self-reported")))

df <- df %>%
  select(-which(sapply(names(df), grepl, pattern = "Non-cancer_illness_code,_self-reported")))

df <- df %>%
  select(-which(sapply(names(df), grepl, pattern = "Treatment/medication_code_diag")))

df <- df %>%
  select(-which(sapply(names(df), grepl, pattern = "Illnesses_of")))
df <- df %>%
  select(-which(sapply(names(df), grepl, pattern = "Medication_for_cholesterol")))

#View(head(df))
dim (df)
```


#removing the negative variables

```{r }

#md.pattern(df_alcohol)
df_rep = df
dim(df_rep)
df_rep = df_rep %>% select (-c(Townsend_deprivation_index_at_recruitment_env.0.0))
#function to replace -3s in the df
replace_neg_three <- function(df) {
  df[df == -3] <- NA
  return(df)
}
df_rep = replace_neg_three (df_rep)



# function to replace the -7 in the df
replace_neg_seven <- function(df) {
  df[df == -7] <- 0
  return(df)
}
df_rep = replace_neg_seven (df_rep)
# function to replace the -7 in the df
replace_neg_one <- function(df) {
  df[df == -1] <- NA
  return(df)
}
df_rep = replace_neg_one (df_rep)

#-6 it varies recorded as yes as it means yes 
df_rep$MealsAndAlc_env.0.0 <- replace(df_rep$MealsAndAlc_env.0.0, 
                                    df_rep$MealsAndAlc_env.0.0
                                    == -6, 1)

# function to replace the -10 in the df
replace_neg_ten <- function(df) {
  df[df == -10] <- 1
  return(df)
}
df_rep = replace_neg_ten (df_rep)
#summary(df_rep)



# # replace all -3 values in all columns except townsend deprivation with NA
# df[, !colnames(df) %in% "Townsend_deprivation_index_at_recruitment_env.0.0"] <- ifelse(df[, !colnames(df) %in% "Townsend_deprivation_index_at_recruitment_env.0.0"] == -3, NA, df[, !colnames(df) %in% "Townsend_deprivation_index_at_recruitment_env.0.0"])


```

# Look at Missing values and creating the ggplot

```{r }
# Finding the percent of missing values for each of the variable
missing_values <- colSums(is.na(df_to_impute))

percent_missing <- round(missing_values/297361 * 100, 2)

# Display the result
missing_data_percent <- data.frame(Percent_of_Missing_data = sort(percent_missing, decreasing = TRUE))

View(missing_data_percent)

# Calculate the percentage of missing values in each variable

missing_pct <- colMeans(is.na(df_rep)) * 100

missing_df <- data.frame(variable = names(missing_pct), missing_pct = missing_pct)

 

# Sort the data frame by missing percentage in decreasing order

missing_df <- missing_df[order(-missing_df$missing_pct), ]

 

# Create the ggplot

ggplot(data = missing_df, aes(x = reorder( variable, missing_pct), y = missing_pct)) +

  geom_bar(stat = "identity", fill = "blue") +

  labs(x = "Variable", y = "Percentage of Missing Values") +

  ggtitle("Percentage of Missing Values by Variable") +

  theme(plot.title = element_text(hjust = 0.5)) +

  geom_hline(yintercept = 30, linetype = "dashed", color = "red") +

  ggtitle("Percentage of missing values in each variable")



```
#Threshold of 30% --> removed some variables except for the NMR ones

```{r }
#   Spirits_weekly_env.0.0 31.07 / Champage_Whitewine_env.0.0 31.01 / RedWine_env.0.0 30.96 / ForitifiedWine_weekly_env.0.0 30.86 / Beer_weekly_env.0.0 30.81

# Removing the variables with more than 30% missingnes

df_rep = df_rep %>% select (-c(  "Age_high_blood_pressure_diagnosed.0.0"
))

# Removing the food variables that were added
df_rep  <- df_rep%>%
  select(-which(sapply(names(df_rep), grepl, pattern = "HLS")))

df_rep$Townsend_deprivation_index_at_recruitment_env.0.0 = df$Townsend_deprivation_index_at_recruitment_env.0.0

#str(df_rep)
#View(head(df_rep))
#str(df_rep)

```



#Converting the variables to their type

```{r }

my_data <- df_rep %>% mutate_at(vars("case", "prevalent_case", "incident_case", "Sex.0.0","Number_of_days/week_of_moderate_physical_activity_10+_minutes_.0.0", "Frequency_of_friend/family_visits.0.0","Time_spent_outdoors_in_winter.0.0","Time_spend_outdoors_in_summer.0.0","Sleep_duration_env.0.0", "Alcohol_env.0.0","Current_tobacco_smoking_env.0.0", "Past_tobacco_smoking.0.0","Smoking_status_env.0.0", "smokers_in_household_env.0.0","diet_variation_env.0.0","MealsAndAlc_env.0.0","Alcohol_intake_versus_10_years_previously_env.0.0","Country_of_birth_env.0.0","Maternal_smoking_around_birth_env.0.0","Smoking_status_env.0.0","Ever_smoked_env.0.0","Ethnic_background_env.0.0","Close_to_major_road_env.0.0"),as.factor) %>%mutate_at(vars("time_to_diagnosis", "TobaccoExposure_home_env.0.0","TobaccoExposure_outside_env.0.0","Diastolic_blood_pressure,_automated_reading_diag.0.0","Diastolic_blood_pressure,_automated_reading_diag.0.1","Systolic_blood_pressure,_automated_reading_diag.0.0", "Systolic_blood_pressure,_automated_reading_diag.0.1","Age_at_recruitment_env.0.0","Traffic_intensity_on_the_nearest_road_env.0.0","Body_mass_index_(BMI)_env.0.0","Traffic_intensity_on_the_nearest_major_road_env.0.0", "Total_traffic_load_on_major_roads_env.0.0"), as.numeric)

str(my_data)
#View(comorbidities)
my_data = my_data %>% select (-c(Diabetes_diagnosis_bio.0.0))

```

 

# Looking at the individuals that have more than 25% missingness 


```{r }

data_rows= my_data %>% select (-c("date_recr", "date_diagnosis","date_death", "time_to_diagnosis", `Date_lost_to_follow-up.0.0`)) %>%
  select(-which(sapply(names(my_data), grepl, pattern = "NMR")))



# calculate thr perfent missing value for each row

na_counts <- apply(is.na(data_rows), 1, sum)


# filter the dataframe to keep only the rows with more NAs in at least 25% of the columns
thresh <- ncol(data_rows) * 0.25  # calculate the threshold for 25% of columns
result <- my_data[na_counts >= thresh, ]


dim(result)


result2 <- anti_join(my_data, result, by = colnames(my_data))
dim(result2)

# # print the result
# nrow(result2)
# 
# missing_values <- colSums(is.na(result2))
# 
# percent_missing <- round(missing_values/297361 * 100, 2)
# 
# # Display the result
# missing_data_percent <- data.frame(Percent_of_Missing_data = sort(percent_missing, decreasing = TRUE))
# View(missing_data_percent)

#str(result2)
rm(data_rows, missing_df)
```
# Merging the comorbidities columns (they are added as binary variables)

```{r }
com_cols <- comorbidities %>% 
  select(starts_with("com"))
com_cols$eid = comorbidities$eid
result2 <- merge(result2 , com_cols, by = "eid")

rm(comorbidities, df_rep, com_cols)
```
# Working with result 2 to prepare the data frame we want to impute

```{r }
df_to_impute = result2

# removing NMR variables and dates from df_to_impute
df_to_impute = df_to_impute%>% select (-c("date_recr", "date_diagnosis","date_death", "time_to_diagnosis", `Date_lost_to_follow-up.0.0`))

df_to_impute = df_to_impute %>%  select(-which(sapply(names(df_to_impute), grepl, pattern = "NMR")))

# removing numerical/factors that will be banded but have NAs
df_to_impute = df_to_impute %>% select (- c(`Number_of_days/week_of_moderate_physical_activity_10+_minutes_.0.0`, Maternal_smoking_around_birth_env.0.0, MealsAndAlc_env.0.0, Alcohol_intake_versus_10_years_previously_env.0.0 , Alcohol_env.0.0, smokers_in_household_env.0.0, Past_tobacco_smoking.0.0, Time_spent_outdoors_in_winter.0.0, Time_spend_outdoors_in_summer.0.0, `Frequency_of_friend/family_visits.0.0`, Sleep_duration_env.0.0, diet_variation_env.0.0, Ethnic_background_env.0.0, Country_of_birth_env.0.0, Smoking_status_env.0.0, Ever_smoked_env.0.0, Current_tobacco_smoking_env.0.0 ))

df_to_impute$com_has_t2d[is.na(df_to_impute$com_has_t2d)] <- 0

write.csv(df_to_impute, file = '/rds/general/project/hda-22-23/live/TDS/Group3_Temporary/imputation/df_all_numeric.csv')

```
# Banding the categorical variables

```{r }
dim(result2)
#str(result2)
result2$physical_activity_level <- cut(as.numeric(result2$`Number_of_days/week_of_moderate_physical_activity_10+_minutes_.0.0`), breaks = c(0, 1, 4, 7), labels = c("Low", "Moderate", "High"))


result2$time_summer_outdoors_level <- cut(as.numeric(result2$Time_spend_outdoors_in_summer.0.0), breaks = c(0, 3, 8, 20), labels = c("Low", "Moderate", "High"))
#plot(result2$time_summer_outdoors_level)

result2$time_winter_outdoors_level <- cut(as.numeric(result2$Time_spent_outdoors_in_winter.0.0), breaks = c(0, 3, 8, 24), labels = c("Low", "Moderate", "High"))


result2$sleep_level <- cut(as.numeric(result2$Sleep_duration_env.0.0), breaks = c(0, 5, 8, 24), labels = c("0-5", "6-8", "9-24"))
#plot(result2$alcohol_level)

result2$alcohol_level <- cut(as.numeric(result2$Alcohol_env.0.0), breaks = c(0, 2, 4, 6), labels = c("0-2", "3-4", "5-6"))

result2$frequency_visits <- cut(as.numeric(result2$`Frequency_of_friend/family_visits.0.0`), breaks = c(1, 2, 3, 5, 7), labels = c("Very frequent", "Once a week", "Rarely", "Never"))

result2$smokers_in_household <- cut(as.numeric(result2$smokers_in_household_env.0.0), breaks = c(0,1, 3), labels = c("Nobody smokes", "At least one person smokes"))

result2$diet_variation <- cut(as.numeric(result2$diet_variation_env.0.0), breaks = c(0,1, 4), labels = c("Never", "Sometimes/Often"))



result2 = result2 %>% select (-c("Number_of_days/week_of_moderate_physical_activity_10+_minutes_.0.0","Time_spend_outdoors_in_summer.0.0" , "Time_spent_outdoors_in_winter.0.0", "Sleep_duration_env.0.0", "Alcohol_env.0.0", `Frequency_of_friend/family_visits.0.0`, smokers_in_household_env.0.0, diet_variation_env.0.0))

result2<- result2 %>% mutate(Ethnic_background = recode(Ethnic_background_env.0.0, `1` = "White",`1001` = "White",`1003` = "White",`1002` = "White",`2` = "Mixed",`2001` = "Mixed",`2002` = "Mixed", `2003` = "Mixed", `2004` = "Mixed",`3001` = "Asian", `3002` = "Asian", `3` = "Asian",`3003` = "Asian", `3004` = "Asian", `5` = "Asian", `4002` = "Black", `4003` = "Black",  `4` = "Black", `4001` = "Black", `6` = "Other"))


```

# Creating the final dataset

```{r }

library(readr)
df_all_numeric_imputed <- read_csv("/rds/general/project/hda-22-23/live/TDS/Group3_Temporary/imputation/df_all_numeric_imputed.csv")

df_all_numeric_imputed = df_all_numeric_imputed %>% select (-c(`Unnamed: 0`))


#Merging the columns from the other dataframe
merged_df <- merge(result2[,c("eid", "date_recr", "date_diagnosis","date_death", "time_to_diagnosis", "Date_lost_to_follow-up.0.0")], df_all_numeric_imputed, by = "eid")

# adding numerical/factors that will be banded but have NAs
merged2_df = merge(merged_df, result2 [,c("eid","physical_activity_level", "Maternal_smoking_around_birth_env.0.0", "MealsAndAlc_env.0.0", "Alcohol_intake_versus_10_years_previously_env.0.0" , "alcohol_level", "smokers_in_household", "Past_tobacco_smoking.0.0", "time_winter_outdoors_level","time_summer_outdoors_level", "frequency_visits", "sleep_level", "diet_variation", "Ethnic_background", "Country_of_birth_env.0.0", "Smoking_status_env.0.0", "Ever_smoked_env.0.0", "Current_tobacco_smoking_env.0.0" ) ], by = "eid") 

# adding the NMR variables
cols_to_select <- grep("NMR", names(result2), value = TRUE)

# merge the two dataframes based on column A
merged2_df <- merge(merged2_df, result2[, c("eid", cols_to_select)], by = "eid")



# we want to make the columns eid and case the first ones#
# select the columns you want to move to the front and create a new dataframe with them

cols_to_move <- c("eid", "case", "prevalent_case", "incident_case", "Sex.0.0","Body_mass_index_(BMI)_env.0.0", "Age_at_recruitment_env.0.0" )

#merged_df <- merge(result2[,c("eid", "case", "prevalent_case", "incident_case", "Sex.0.0","Body_mass_index_(BMI)_env.0.0", "Age_at_recruitment_env.0.0")], merged2_df, by = "eid", "case", "prevalent_case", "incident_case", "Sex.0.0","Body_mass_index_(BMI)_env.0.0", "Age_at_recruitment_env.0.0")
merged_df <- merged2_df[, c(cols_to_move, setdiff(names(merged2_df), cols_to_move))]




```

#Removing variables based on the correlation heatmap and calculating the blood pressure
 
```{r }


# calculating the final Blood pressure
cleaned_df <- merged_df%>%
  mutate(MAP = `Diastolic_blood_pressure,_automated_reading_diag.0.0` +
           (`Systolic_blood_pressure,_automated_reading_diag.0.0` -
              `Diastolic_blood_pressure,_automated_reading_diag.0.0`)/3)


# so we kept the variables that are measured in 2010, for the percentage buffer we considered for water and natural environment 1000 as we need to consider a larger space, thus removed the 300m. 
# For domestic garden we kep the 300 m because 1000m would be less specifc
cleaned_df = cleaned_df %>% select(-c( `Greenspace_percentage,_buffer_300m_env.0.0`, `Greenspace_percentage,_buffer_1000m_env.0.0`, `Water_percentage,_buffer_300m_env.0.0`, `Greenspace_percentage,_buffer_300m_env.0.0`,`Domestic_garden_percentage,_buffer_1000m_env.0.0`, `Natural_environment_percentage,_buffer_300m_env.0.0`, `Water_percentage,_buffer_300m_env.0.0`))

cleaned_df = cleaned_df %>% select (-c(`Mean_corpuscular_volume_bio.0.0`,`Mean_platelet_(thrombocyte)_volume_bio.0.0`, Nucleated_red_blood_cell_count_bio.0.0,Mean_reticulocyte_volume_bio.0.0, Mean_sphered_cell_volume_bio.0.0, High_light_scatter_reticulocyte_count_bio.0.0)) 

cleaned_df = cleaned_df %>% select (-c(bio_blood_Apolipoprotein_A.0.0, bio_blood_Apolipoprotein_B_bio.0.0, bio_blood_Cholesterol.0.0, bio_blood_Direct_bilirubin.0.0, bio_blood_Aspartate_aminotransferase.0.0))

cleaned_df = cleaned_df  %>% select(-c("Systolic_blood_pressure,_automated_reading_diag.0.1", "Diastolic_blood_pressure,_automated_reading_diag.0.1", "Diastolic_blood_pressure,_automated_reading_diag.0.0", "Systolic_blood_pressure,_automated_reading_diag.0.0"))
                                   

cleaned_df = cleaned_df  %>% select(-c(Average_evening_sound_level_of_noise_pollution_env.0.0, Average_daytime_sound_level_of_noise_pollution_env.0.0, Sum_of_road_length_of_major_roads_within_100m_env.0.0))

cleaned_df = cleaned_df  %>% select(-c(Inverse_distance_to_the_nearest_major_road_env.0.0, Traffic_intensity_on_the_nearest_major_road_env.0.0))



```


# converting the NAs in factors to unknown
```{r }

#str(cleaned_df)


cleaned2_df <- cleaned_df %>%
mutate_if(is.factor, function(x) { ifelse(is.na(x), "Unknown", x)} )

# recpnvert to the correct data type
cleaned2_df <- cleaned2_df %>% mutate_at(vars("case", "prevalent_case", "incident_case", "Sex.0.0", "frequency_visits","Current_tobacco_smoking_env.0.0", "Past_tobacco_smoking.0.0","smokers_in_household","diet_variation","MealsAndAlc_env.0.0","Alcohol_intake_versus_10_years_previously_env.0.0","Country_of_birth_env.0.0","Maternal_smoking_around_birth_env.0.0","Smoking_status_env.0.0","Ever_smoked_env.0.0","Close_to_major_road_env.0.0", "physical_activity_level", "time_summer_outdoors_level", "time_winter_outdoors_level", "sleep_level", "alcohol_level", "Ethnic_background","com_has_AF" , "com_has_migraine","com_has_t1d","com_has_t2d"  ,"com_has_ckd","com_has_RA","com_has_sle","com_has_psychiatric" ,"com_has_ED" , "com_herit_CAD" ,"com_antipsychotic","com_steroid" ,"com_lipid_lowering" ,"com_bp" ),as.factor)



cleaned_df = cleaned2_df
table(result2$com_bp)
#table(result2$com_lipid_lowering, result2$eid)
part = result2$eid [ ]

participants <- comorbidities[comorbidities$com_lipid_lowering == "1", "eid"]
participants = lapply (participants, as.numeric)
count <- sum(participants %in% result2$eid)
count


```
